---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

### What does exdex do?

The `exdex` package uses threshold inter-exceedance times based on the K-gaps 
model and the method described in Northrop (2015) that is based on block maxima 
to estimate the extremal index. The extremal index measures how strong is the dependence between the threshold exceedances. 

## Installation

You can install exdex from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("ConstantinosChr/exdex")
```

## A simple example

For the K-gaps method:

```{r}
thresh <- quantile(newlyn, probs = 0.90)
k_res <- kgaps_mle(newlyn, thresh, conf = 95)
k_res$theta_mle
k_res$theta_se
k_res$theta_ci
```

For the block maxima method:

```{r}
spm_mle(newlyn, 20)
spm_mle(newlyn, 20, sliding = FALSE)
```

## Vignette

See \code{vignette("exdex-vignette", package = "exdex")} for an overview of the package.
